---
title: "Video analysis for urchin deterrence experiment"
author: "Amelia Ritger"
date: "10/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

### Load packages and read in data
```{r}
library(tidyverse)
library(ggplot2)
library(GGally)
library(janitor)

data <- read.csv("data/raw.csv")
```

### Tidy data
```{r}
# my own OCD, fix "Trial" issue
colnames(data)[colnames(data)=="ï..Trial.number"] <- "Trial"
# Clean names for R
data <- data %>%
  clean_names()
# remove trials without videos
data <- subset(data, !tank_number>6 & !trial==46)
# remove instances where urchin never came in contact with tile during video (regardless of kelp consumption after video ended)
data <- subset(data, !data$urchin_deterred_during_video=="urchin never tried")
# remove instances where urchin never came in contact with tile during video (but keep instances where kelp consumption occurred after video ended because otherwise n=3 for control treatment)
#data <- subset(data, !urchin_sucessfully_deterred_during_video=="Urchin never moved")
# To compare across treatments, presence/absence of Corynactis
data$corynactis_binary <- ifelse(data$treatment_number>1, "present", "absent")
# Create binary variable: urchin crossed (yes/no)
data$cross_binary <- ifelse(data$times_crossing_cory>0, 1, 0)
# Create variables: kelp area consumed, percent kelp consumed
data <- data %>%
  mutate(area_consumed=(kelp_before_cm_2-kelp_after_cm_2)) %>%
  mutate(percent_consumed=(area_consumed/kelp_before_cm_2))
# Now filter out values where urchins didn't actually consume kelp
data$area_corrected <- ifelse(data$kelp_visibly_consumed=="no", 0, data$area_consumed)
data$percent_corrected <- ifelse(data$kelp_visibly_consumed=="no", 0, data$percent_consumed)
# To compare consumption (yes, no) across treatments
data$consumption_binary <- ifelse(data$percent_corrected>0, 1, 0)
# Create variable for binary deterrence (was urchin ever deterred)
data$deter_binary <- ifelse(data$times_deterred_from_cory>0, 1, 0)
# Average urchin body weight before and after trials
data <- data %>%
  mutate(urchin_avg_g = rowMeans(cbind(urchin_weight_before_g,urchin_weight_after_g)))
# Standardize "instances of action" across trials because different video lengths
data <- data %>%
  mutate(contact_per_hr=((times_in_contact_with_cory/total_video_time_min)*60)) %>%
  mutate(deter_per_hr=((times_deterred_from_cory/total_video_time_min)*60)) %>%
  mutate(cross_per_hr=((times_crossing_cory/total_video_time_min)*60)) %>%
  #mutate(deter_per_hr = replace(deter_per_hr, deter_per_hr == 0, 0.000001)) %>%
  mutate(percent_deter=(deter_per_hr/contact_per_hr)) %>%
  mutate(percent_cross=(cross_per_hr/contact_per_hr))
# Clean up data frame, only keep variables you want to work with
data <- data %>%
  dplyr::select(julian_date, kelp_location, urchin_size_mm, urchin_avg_g, treatment, corynactis_binary, consumption_binary, area_corrected, percent_corrected, time_to_cross_cory_first_min, percent_time_alone, percent_time_with_kelp, when_was_kelp_consumed, deter_binary, times_in_contact_with_cory, times_deterred_from_cory, urchin_deterred_during_video, urchin_sucessfully_deterred_during_video, contact_per_hr, deter_per_hr, cross_per_hr, percent_deter, percent_cross)
```


### Visualize data
```{r, fig.show="hide"}
#look for correlations among variables
ggpairs(data[, c("julian_date", "kelp_location", "urchin_size_mm", "urchin_avg_g", "treatment", "corynactis_binary", "consumption_binary", "area_corrected", "percent_corrected", "time_to_cross_cory_first_min", "percent_time_alone", "percent_time_with_kelp", "when_was_kelp_consumed", "urchin_deterred_during_video", "urchin_sucessfully_deterred_during_video", "contact_per_hr", "deter_per_hr", "cross_per_hr", "percent_deter", "percent_cross")])

#look at treatment with everything
#look at corynactis_binary with everything
#come back to this when making final conclusions, because lots of the Ys are highly correlated with eachother (obviously) so you want to avoid being  repetitive by commenting on different relationships with correlated variables
# % time with kelp is - corr with % time with alone (-.98) and both are corr with # deterrence (0.58, -0.57)
# number of contacts with tile is + corr with number deterrence events (0.81) and number crosses (0.38)
#interesting things to look at separately - # deterrence, # crosses, time to cross tile
```

## How does presence of Corynactis affect "outcome" (see data frame)?
```{r}
X <- as.factor(data$corynactis_binary)
outcome <- factor(as.factor(data$urchin_sucessfully_deterred_during_video))

t1 <- table(outcome, X) #because data is binary and sample size is small (controls n = 5), use N-1 Two Proportion Test (below)
prop.test(t1, correct=TRUE) # correct=TRUE tells R to correct for small sample size p = 0.0393
fisher.test(t1) #fisher's exact test is useful when sample sizes are particularly small... p=0.0505
```

Outcome: Urchins always crossed tile when it didn't have Corynactis significantly more than expected


## How does treatment type affect "outcome" (see data frame)?
```{r, fig.show="hide"}
library(corrplot)
X <- as.factor(data$treatment)
t1 <-table(X, outcome)
t1 = t1[-1,] #because "Control" is driving whole relationship
cs<-chisq.test(t1)
cs$p.value #p = 0.2054, outcome and treatment are independent
corrplot(cs$residuals, is.cor = FALSE) #visualize pearsons residuals; (blue color means positively associated with variables, red color means negatively associated with variables); controls driving a lot of this trend, pink and red also important
fisher.test(t1)
```

Outcome: Treatment type doesn't affect outcome

## How does presence of Corynactis affect whether or not urchin was deterred (regardless of ultimate outcome)? 
```{r}
X <- as.factor(data$corynactis_binary)
deter <- as.factor(data$deter_binary)

t1 <- table(X, deter)
prop.test(t1, correct=TRUE) #two proportion test p = 0.01524
fisher.test(t1) # fisher's test p = 0.0067
```

Outcome: Urchin was never detered from tile when Corynactis was absent, urchins were likely to be deterred when Corynactis present

## How does type of Corynactis affect whether or not urchin was deterred (regardless of ultimate outcome)?
```{r}
X <- as.factor(data$treatment)

t1 <- table(X, deter)
t1 = t1[-1,] # because "Control" is driving the whole relationship
prop.test(t1, correct=TRUE) #p=0.8174
fisher.test(t1)
```

Outcome: Treatment type definitely doesn't affect whether or not an urchin was deterred at least 1x

## Time with kelp: correlation with amount of kelp consumed? (and is there an urchin size effect?)
```{r}
time <- data$percent_time_with_kelp
area <- data$area_corrected
size <- data$urchin_avg_g
size_mm <- data$urchin_size_mm
hist(time)
hist(area)
hist(size)
hist(size_mm)

plot(area~size_mm)
abline(lm(area~size_mm))
plot(time~size_mm)
abline(lm(time~size_mm))
plot(area~time)
abline(lm(area~time))

m1<- lm(area~time+size+(time:size))
summary(m1)
summary(lm(area~size)) # >> follow up on this, especially fixed effects model
```

Outcome: larger urchins weren't eating more kelp, larger urchins weren't spending more time with kelp, if anything they spent less time with kelp, more time spent with kelp = more kelp consumed

## Presence of Corynactis: effect on time to cross tile (urchin size effect?)
```{r, fig.show="hide"}
cross <- data %>%
  dplyr::select(treatment, corynactis_binary, time_to_cross_cory_first_min, urchin_avg_g, urchin_size_mm) %>%
  filter(!is.na(time_to_cross_cory_first_min)) %>%
  filter(time_to_cross_cory_first_min<100) #remove this one point, it's a huge outlier
Y <- log10(cross$time_to_cross_cory_first_min) #to meet assumptions of normality
X <- as.factor(cross$corynactis_binary)
hist(Y)
library(car)
library(data.table)
qqPlot(Y) #good enough

## importantly... is there an effect of urchin size?
plot(Y~cross$urchin_size_mm)
abline(lm(Y~cross$urchin_size_mm))
plot(Y~cross$urchin_avg_g)
abline(lm(Y~cross$urchin_avg_g))
summary(lm(Y~cross$urchin_avg_g)) #p=0.8554
```

Definitely no effect of urchin size on time it takes them to cross the tile

```{r}
plot(Y~X)
t.test(Y~X) #p=0.039
```

When Corynactis present, time to cross tile was significantly longer

## Type of Corynactis: effect on time to cross tile
```{r}
X <-  as.factor(cross$treatment)
plot(Y~X)
m1 <- lm(Y~X)
a1 <- aov(Y~X)
summary(a1) #p=0.0034
TukeyHSD(a1)
```

Outcome: All treatment types are significantly different from control, but there is no signficant difference amongst colors

## Presence of Corynactis: effect on % of contacts that resulted in deterrence ("percent_deter")? (urchin size effect?)
```{r, fig.show="hide"}
data$deter_corrected <- ifelse(data$percent_deter=="NaN", 0, data$percent_deter)
deter <- data %>%
  dplyr::select(treatment, corynactis_binary, deter_corrected, urchin_avg_g, urchin_size_mm)
ggplot(deter, aes(x=corynactis_binary, y=deter_corrected))+
  geom_jitter()

Y <- deter$deter_corrected
hist(Y)
qqPlot(Y)
# >> there are lots of zeros, need to correct for that or account for that in the analysis
```
Interpretation:
Mu represents the mean of the beta distribution part of the model for the interval (0,1)
Sigma represents the precision (or shape) of the beta distribution part of the model
Nu represents the Bernoulli distribution part of the model and probability of observing zeros

## Type of Corynactis: effect on % of contacts that resulted in deterrence ("percent_deter")? (urchin size effect?)
```{r}
ggplot(deter, aes(x=treatment, y=deter_corrected)) +
  geom_boxplot() +
  geom_jitter(width=0.1)
```

Maybe there is a difference between orange and pink/red?

```{r}
Y <- deter$deter_corrected
X <- as.factor(deter$treatment)
hist(Y)

deter <- data %>%
  mutate(success=data$times_deterred_from_cory) %>%
  mutate(failure=data$times_in_contact_with_cory-data$times_deterred_from_cory) %>%
  dplyr::select(treatment,success,failure)

#treat data like Bernoulli distribution, each contact by urchin can result in either success or failure
m1 <- glm(cbind(success, failure) ~ treatment, family = binomial , data = deter)
summary(m1)

#now treat data like zero- and one-inflated beta distribution
library(gamlss)
b1 <- gamlss(Y~X,  family = BEINF, trace = F) #BEINF because lots of zero and some one values are present in beta distribution
summary(b1)

#not really sure what this can be used for
#means_b1 <- lpred(b1, type='response', what='mu', se.fit=T)
#df_fit <- data.frame(CORY = data$corynactis_binary, M = means_b1$fit, SE = means_b1$se.fit)
#ggplot(df_fit, aes(CORY, M)) + geom_pointrange(aes(ymin=M-SE, ymax=M+SE)) + 
#  labs(x="Corynactis presence/absence",y="% contacts that resulted in deterrence") +
#  scale_y_continuous(labels=scales::percent)
```

Nope, there is no significant difference.

### And finally, plotting some possibly correlated variables

```{r}
ggplot(data, aes(y=deter_per_hr, x=kelp_location)) +
  geom_boxplot()
```
No effect

```{r}
ggplot(data, aes(y=percent_corrected, x=kelp_location)) +
  geom_boxplot()
```
No effect

```{r}
ggplot(data, aes(y=percent_time_alone, x=urchin_avg_g)) +
  geom_point()
```
Not really an effect
  
```{r}
ggplot(data, aes(x=julian_date, y=urchin_avg_g)) +
  geom_point()
```
As expected, urchins were either big or small later on in experiment as I ran out of urchins

```{r}
ggplot(data, aes(x=treatment, y=urchin_avg_g)) +
  geom_boxplot()
```
No real correlation between treatment and urchin size used