---
title: "Video analysis for urchin deterrence experiment"
author: "Amelia Ritger"
date: "10/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width="400px", out.height="300px")
```

### Load packages, load data, tidy data

##### *magic happening behind the scenes*
```{r,echo=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(GGally)
library(janitor)
library(here)

data <- read_csv(here::here("data", "raw.csv"))

# my own OCD, fix "Trial" issue
colnames(data)[colnames(data)=="ï..Trial.number"] <- "Trial"
# Clean names for R
data <- data %>%
  clean_names()
# remove trials without videos
data <- subset(data, !tank_number>6 & !trial_number==46)
# remove instances where urchin never came in contact with tile during video (regardless of kelp consumption after video ended)
data <- subset(data, !data$urchin_deterred_during_video=="urchin never tried")
# remove instances where urchin never came in contact with tile during video (but keep instances where kelp consumption occurred after video ended because otherwise n=3 for control treatment)
#data <- subset(data, !urchin_sucessfully_deterred_during_video=="Urchin never moved")
# To compare across treatments, presence/absence of Corynactis
data$corynactis_binary <- ifelse(data$treatment_number>1, "present", "absent")
# Create binary variable: urchin crossed (yes/no)
data$cross_binary <- ifelse(data$times_crossing_cory>0, 1, 0)
# Create variables: kelp area consumed, percent kelp consumed
data <- data %>%
  mutate(area_consumed=(kelp_before_cm_2-kelp_after_cm_2)) %>%
  mutate(percent_consumed=(area_consumed/kelp_before_cm_2))
# Now filter out values where urchins didn't actually consume kelp
data$area_corrected <- ifelse(data$kelp_visibly_consumed=="no", 0, data$area_consumed)
data$percent_corrected <- ifelse(data$kelp_visibly_consumed=="no", 0, data$percent_consumed)
# To compare consumption (yes, no) across treatments
data$consumption_binary <- ifelse(data$percent_corrected>0, 1, 0)
# Create variable for binary deterrence (was urchin ever deterred)
data$deter_binary <- ifelse(data$times_deterred_from_cory>0, 1, 0)
# Average urchin body weight before and after trials
data <- data %>%
  mutate(urchin_avg_g = rowMeans(cbind(urchin_weight_before_g,urchin_weight_after_g)))
# Standardize "instances of action" across trials because different video lengths
data <- data %>%
  mutate(contact_per_hr=((times_in_contact_with_cory/total_video_time_min)*60)) %>%
  mutate(deter_per_hr=((times_deterred_from_cory/total_video_time_min)*60)) %>%
  mutate(cross_per_hr=((times_crossing_cory/total_video_time_min)*60)) %>%
  #mutate(deter_per_hr = replace(deter_per_hr, deter_per_hr == 0, 0.000001)) %>%
  mutate(percent_deter=(deter_per_hr/contact_per_hr)) %>%
  mutate(percent_cross=(cross_per_hr/contact_per_hr))
# Clean up data frame, only keep variables you want to work with
data <- data %>%
  dplyr::select(julian_date, kelp_location, urchin_size_mm, urchin_avg_g, treatment, corynactis_binary, consumption_binary, area_corrected, percent_corrected, time_to_cross_cory_first_min, percent_time_alone, percent_time_with_kelp, when_was_kelp_consumed, deter_binary, times_in_contact_with_cory, times_deterred_from_cory, urchin_deterred_during_video, urchin_sucessfully_deterred_during_video, contact_per_hr, deter_per_hr, cross_per_hr, percent_deter, percent_cross)
```

### Some prelimiary analysis: plotting some possibly correlated variables

```{r, echo=FALSE}
ggplot(data, aes(y=deter_per_hr, x=kelp_location)) +
  geom_boxplot()
```

No correlation between where kelp was placed in tank and deterrences per hour

```{r, echo=FALSE}
ggplot(data, aes(y=percent_corrected, x=kelp_location)) +
  geom_boxplot()
```

No correlation between where kelp was placed in tank and percent of kelp consumed


```{r, echo=FALSE}
ggplot(data, aes(y=percent_time_alone, x=urchin_avg_g)) +
  geom_point()
```

No real correlation between urchin size and percent of time urchin spent alone during trial

```{r, echo=FALSE}
ggplot(data, aes(x=julian_date, y=urchin_avg_g)) +
  geom_point()
```

As expected, urchins were either big or small later on in experiment as I ran out of urchins

```{r, echo=FALSE}
ggplot(data, aes(x=treatment, y=urchin_avg_g)) +
  geom_boxplot()
```

No real correlation between treatment and urchin size used

### Is there any correlation with amount of kelp consumed and urchin size?
```{r, echo=FALSE, message=FALSE, fig.show="hide"}
time <- data$percent_time_with_kelp
area <- data$area_corrected
size <- data$urchin_avg_g
size_mm <- data$urchin_size_mm
hist(time)
hist(area)
hist(size)
hist(size_mm)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot(area~size_mm)
abline(lm(area~size_mm))
```

Urchin size had no correlation with amount of kelp consumed

```{r, echo=FALSE, message=FALSE, fig.show="hide"}
plot(time~size_mm)
abline(lm(time~size_mm))
```


```{r, echo=FALSE, message=FALSE}
plot(area~time)
abline(lm(area~time))
```

But the more time spent with kelp >> more kelp consumed (no surprise) 

How might these variables interact with kelp consumption?
```{r, echo=FALSE, message=FALSE}
m1<- lm(area~time+size+(time:size))
summary(m1)
#summary(lm(area~size)) # >> follow up on this, especially fixed effects model
```
It doesn't seem like, statistically, there are any effects or interacting effects on area of kelp consumed

So: larger urchins weren't eating more kelp, larger urchins weren't spending more time with kelp, if anything they spent less time with kelp, and (obviously) more time spent with kelp = more kelp consumed

### Is there an effect of urchin size on time to cross tile?
```{r, include=FALSE}
cross <- data %>%
  dplyr::select(treatment, corynactis_binary, time_to_cross_cory_first_min, urchin_avg_g, urchin_size_mm) %>%
  filter(!is.na(time_to_cross_cory_first_min)) %>%
  filter(time_to_cross_cory_first_min<50) #remove this one point, it's a huge outlier
Y <- log10(cross$time_to_cross_cory_first_min) #to meet assumptions of normality
X <- as.factor(cross$corynactis_binary)
hist(Y)
library(car)
library(data.table)
qqPlot(Y) #good enough
```
Time to cross tile (Y) has been log transformed to meet assumptions of normality

```{r, echo=FALSE}
##Is there an effect of urchin size?
plot(Y~cross$urchin_size_mm)
abline(lm(Y~cross$urchin_size_mm))
plot(Y~cross$urchin_avg_g)
abline(lm(Y~cross$urchin_avg_g))
summary(lm(Y~cross$urchin_avg_g)) #p=0.8554
```
Definitely no effect of urchin size on time it takes them to cross the tile

# Now onto the meat of the analysis

## How does presence of Corynactis affect "outcome"?

#### __Outcome options:__

- Cory was beast = Urchin never crossed over Corynactis tile

- Cory was close = Urchin was deterred from Corynactis tile at least once but ultimately crossed over tile

- Uchin was beast = Urchin was never deterred from Corynactis tile

#### Because data is binary and sample size is small (control n = 3), use N-1 Two Proportion Test or Fisher's Exact Test
```{r, echo=FALSE, message=FALSE, warning=FALSE}
X <- as.factor(data$corynactis_binary)
outcome <- factor(as.factor(data$urchin_sucessfully_deterred_during_video))

t1 <- table(outcome, X)
print(t1)
prop.test(t1, correct=TRUE) # correct=TRUE tells R to correct for small sample size p = 0.0393 for n=5
fisher.test(t1) #fisher's exact test is useful when sample sizes are particularly small... p=0.0505 for n=5
```
Urchins always crossed tile when it didn't have Corynactis, but it wasn't statistically significant (probably because of low control sample size)


## How does treatment type affect "outcome"?
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hide"}
library(corrplot)
X <- as.factor(data$treatment)
t1 <-table(X, outcome)
t1 = t1[-1,] #because "Control" is driving whole relationship
cs<-chisq.test(t1)
print(cs)
corrplot(cs$residuals, is.cor = FALSE) #visualize pearsons residuals; (blue color means positively associated with variables, red color means negatively associated with variables); controls driving a lot of this trend, pink and red also important
fisher.test(t1)
```
Different color morphs definitely did not result in different outcomes

## How does presence of Corynactis affect whether or not urchin was deterred (regardless of ultimate outcome)? 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
X <- as.factor(data$corynactis_binary)
deter <- as.factor(data$deter_binary)

t1 <- table(X, deter)
print(t1)
prop.test(t1, correct=TRUE) #two proportion test p = 0.01524 when n=5 for controls
fisher.test(t1) # fisher's test p = 0.0067 when n=5 for controls
```
Urchin was never detered from tile when Corynactis was absent, urchins were likely to be deterred when Corynactis present, but this is not really statistically significant (it depends which test you use)

## How does type of Corynactis affect whether or not urchin was deterred (regardless of ultimate outcome)?
```{r, echo=FALSE, message=FALSE, warning=FALSE}
X <- as.factor(data$treatment)

t1 <- table(X, deter)
t1 = t1[-1,] # because "Control" is driving the whole relationship
prop.test(t1, correct=TRUE) #p=0.8174
fisher.test(t1)
```
Treatment type does not affect whether or not an urchin was deterred at least once

## Is there an effect of Corynactis (yes/no) on time to cross tile?
```{r, echo=FALSE}
ggplot(cross, aes(x=corynactis_binary, y=time_to_cross_cory_first_min)) +
  geom_boxplot(aes(fill=corynactis_binary), outlier.shape=NA, show.legend=FALSE) +
  geom_jitter(alpha=0.4, width=0.1, shape=1, stroke = 1.5) +
  scale_fill_manual(values=c("cadetblue4", "coral")) +
  labs(x="", y="Time for urchin to cross tile (min)") +
  scale_x_discrete(labels = c('Corynactis absent','Corynactis present')) +
  scale_y_continuous(limits=c(0,50),expand = c(0,0), breaks=c(0,10,20,30,40,50)) +
  theme(panel.grid.major.y  = element_line(color = "grey90"),
      panel.background = element_rect(fill = "white"),
      panel.border = element_rect(color = "white", fill = NA),
      axis.line = element_line(color = "black"),
      axis.ticks = element_line(color = "black"),
      axis.text = element_text(color = "black", size=12),
      axis.title.y = element_text(size=14, vjust=4),
      plot.margin=unit(c(0.5,0.2,-0.3,0.5), "cm"))

ggsave(here::here("figures", "presence by time to cross tile.png"), width=6, height=4.5)
```
<< This is a potential plot for WSN poster

```{r, echo=FALSE}
t.test(log10(cross$time_to_cross_cory_first_min)~as.factor(cross$corynactis_binary)) #p=0.039
```
When Corynactis were present, the time it took an urchin to cross the tile was significantly longer

## Is there an effect of treatment on time to cross tile?
```{r, echo=FALSE}
plot(log10(cross$time_to_cross_cory_first_min)~as.factor(cross$treatment))
m1 <- lm(log10(cross$time_to_cross_cory_first_min)~as.factor(cross$treatment))
a1 <- aov(log10(cross$time_to_cross_cory_first_min)~as.factor(cross$treatment))
summary(a1) #p=0.0034
TukeyHSD(a1)
```
All experimental treatments are significantly different from control treatment, but there is no signficant difference amongst color morphs

## Presence of Corynactis: effect on % of contacts that resulted in deterrence?
```{r, echo=FALSE}
data$deter_corrected <- ifelse(data$percent_deter=="NaN", 0, data$percent_deter)
data$deter_trans <- asin(sqrt(data$deter_corrected))
data$deter_log <- log(data$deter_corrected/(1-data$deter_corrected))
hist(data$deter_log) #hot damn that's normal but also tons of -Inf
qqPlot(data$deter_trans)
qqPlot(data$deter_corrected) #no difference between arcsin sqrt transformed and non transformed

deter <- data %>%
  dplyr::select(treatment, corynactis_binary, deter_corrected)

ggplot(deter, aes(x=corynactis_binary, y=deter_corrected))+
  geom_boxplot() +
  geom_jitter(size=1.5,alpha=0.9, width=0.1) +
  labs(x="Treatment", y="% contacts that resulted in urchin deterrence") +
  theme_bw()

Y <- deter$deter_corrected
hist(Y)
qqPlot(Y)
```
### There are many zeros that have not been corrected for in the analysis
#### AND this is not really a different question from the "presence affecting outcome" question

## Type of Corynactis: effect on % of contacts that resulted in deterrence?
```{r, echo=FALSE}
ggplot(deter, aes(x=treatment, y=deter_corrected)) +
  geom_boxplot() +
  geom_jitter(width=0.1)
```

Maybe there is a difference between orange and pink/red?

```{r, echo=FALSE, message=FALSE}
#remove control
#deter <- subset(deter, !deter$treatment=="Control") has even less of an effect
a1 <- aov(deter$deter_corrected~deter$treatment)
summary(a1) #p=0.11
TukeyHSD(a1) #non significant but some difference between pink/red and control (but certainly no difference between color morphs)
```
No, definitely no difference.

Plot means + confidence intervals of the groups
```{r}
library(Rmisc)
deterred <- summarySE(deter, measurevar="deter_corrected", groupvars="treatment")
mean <- deterred$deter_corrected
ci <- deterred$ci

# ggplot(deter, aes(x=treatment, y=deter_corrected, color=treatment)) +
#   geom_jitter(size=4, alpha=0.5, width=0.1, show.legend=FALSE) +
#   scale_color_manual(values=c("cadetblue4", "orange", "pink", "red")) +
#   geom_segment(data=deterred, aes(x=c(0.5,1.5,2.5,3.5), xend=c(1.5,2.5,3.5,4.5), y=mean, yend=mean), size=3, show.legend=FALSE) +
#   geom_segment(data=deterred, aes(x=c(1,2,3,4), xend=c(1,2,3,4), y=mean+ci, yend=mean-ci), size=1.5, show.legend=FALSE, alpha=0.5) +
#   geom_segment(x=2, xend=2, y=0.2763889+0.2954399, yend=0.2763889-0.2954399, size=2, color="orange", alpha=0.2)+
#   labs(x="Treatment", y="Proportion of contacts resulting in deterrence") +
#   #scale_x_discrete(labels = c('Corynactis absent','Corynactis present')) +
#   scale_y_continuous(limits=c(0,1.05),expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
#   coord_flip() +
#   theme(panel.grid.major.x  = element_line(color = "grey90"),
#       panel.background = element_rect(fill = "white"),
#       panel.border = element_rect(color = "white", fill = NA),
#       axis.line = element_line(color = "black"),
#       axis.ticks = element_line(color = "black"),
#       axis.text = element_text(color = "black", size=12),
#       axis.title.y = element_text(vjust=4, size=14),
#       axis.title.x = element_text(size=14, vjust=-1.5),
#       plot.margin=unit(c(0.2,0.2,0.5,0.5), "cm"))
```

```{r}
ggplot(deterred, aes(x=treatment, y=deter_corrected, color=treatment)) +
  geom_point(aes(color=treatment), shape=124, size=15, show.legend=FALSE) +
  geom_segment(aes(y=deter_corrected-ci, x=treatment, yend=deter_corrected+ci, xend=treatment), show.legend=FALSE) +
  geom_segment(x=2, xend=2, y=0.2763889+0.2954399, yend=0.2763889-0.2954399, color="orange")+
  scale_color_manual(values=c("cadetblue4", "orange", "pink", "red")) +
  #geom_segment(aes(y=deter_corrected-ci, x=treatment, yend=deter_corrected+ci, xend=treatment), show.legend=FALSE) +
  geom_jitter(data=deter, size=2, alpha=0.7, width=0.1, show.legend=FALSE) +
  labs(x="Treatment", y="Proportion of contacts with tile resulting in urchin deterrence") +
  scale_y_continuous(limits=c(-0.01,1.05),expand = c(0,0), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  coord_flip() +
  theme(panel.grid.major.x  = element_line(color = "grey90"),
      panel.background = element_rect(fill = "white"),
      panel.border = element_rect(color = "white", fill = NA),
      axis.line = element_line(color = "black"),
      axis.ticks = element_line(color = "black"),
      axis.text = element_text(color = "black", size=12),
      axis.title.y = element_text(vjust=4, size=14),
      axis.title.x = element_text(size=14, vjust=-1.5),
      plot.margin=unit(c(0.2,0.2,0.5,0.5), "cm"))

ggsave(here::here("figures", "treatment by percent deterrence.png"), width=6, height=4.5)
```


```{r, echo=FALSE, message=FALSE, fig.show="hide"}
Y <- deter$deter_corrected
X <- as.factor(deter$treatment)
hist(Y)
deter <- data %>%
  mutate(success=data$times_deterred_from_cory) %>%
  mutate(failure=data$times_in_contact_with_cory-data$times_deterred_from_cory) %>%
  dplyr::select(treatment,success,failure)
```

### Treat data like Bernoulli: each contact by urchin can result in either success or failure. Also treat data as a beta distribution with zero and one inflation.
```{r, echo=FALSE, message=FALSE}
m1 <- glm(cbind(success, failure) ~ treatment, family = binomial , data = deter)
summary(m1)
#now treat data like zero- and one-inflated beta distribution
library(gamlss)
b1 <- gamlss(Y~X,  family = BEINF, trace = F) #BEINF because lots of zero and some one values are present in beta distribution
summary(b1)
#not really sure what this can be used for
#means_b1 <- lpred(b1, type='response', what='mu', se.fit=T)
#df_fit <- data.frame(CORY = data$corynactis_binary, M = means_b1$fit, SE = means_b1$se.fit)
#ggplot(df_fit, aes(CORY, M)) + geom_pointrange(aes(ymin=M-SE, ymax=M+SE)) + 
#  labs(x="Corynactis presence/absence",y="% contacts that resulted in deterrence") +
#  scale_y_continuous(labels=scales::percent)
```
###### Interpretation:

###### - Mu represents the mean of the beta distribution part of the model for the interval (0,1)

###### - Sigma represents the precision (or shape) of the beta distribution part of the model

###### - Nu represents the Bernoulli distribution part of the model and probability of observing zeros

Nope, there is no significant difference between treatments and % of contacts that resulted in deterrence


# So, in conclusion:
- #### 100% of the time, urchins were never deterred by tiles without Corynactis (but it wasn’t statistically significant)
- #### 37.5% of the time, urchins were never deterred by tiles with Corynactis
- #### There was no significant difference between color morphs affecting the "outcome"
- #### Corynactis caused urchins to take longer to cross the tile (but there was no effect of color morph)
- #### Larger urchins did not spend more time with kelp or eat more kelp
- #### Larger urchins did not cross the Corynactis tile more quickly (or slowly) than smaller urchins

### AKA: Corynactis had an effect on urchin behavior as we measured it with GoPros, but there wasn't necessarily a difference between color morphs on urchin behavior